Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors That Have Mixed During Cooling

### 2019-10-27 paper/ms.tex
Tiny formating issue- removed a space.
Took at the blue

{\color{blue}
\subsection{Proposed Progenitor Settings}\label{sec:progenitors}

@@

In the DDT paradigm, the duration of the deflagration phase sets
the amount of expansion of the star prior to the bulk of burning,
which is critical to the composition of material synthesized in the 
explosion. Also, early burning during the deflagration phase is at 
high enough densities that the effects of electron capture are 
significant and similarly influence the composition of the material
synthesized in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
}

@@

{\color{blue} 
Discerning the role played by hybrid progenitors in the global gamut of
thermonuclear supernovae is the goal of the study we present in this paper. Our simulations
provide the yields of \Ni{56} in the supernova explosions. Because
the radioactive decay of \Ni{56} powers the light curve of an event, 
the yield of \Ni{56} serves as a proxy for brightness of the event so
our estimates allows us to discern trends in brightness. Our study thus 
can offer insight into the relative brightness of explosions from these
progenitors as well as the overall scatter in the brightness
of observed events.

Our simulation methodology is described in the next section and
our results include a comparison of nucleosynthetic explosion yields between these
new mixed hybrid C-O-Ne progenitors and traditional C-O progenitors
in the section following the methodology. Our results show systematic
differences in the yield of \Ni{56} between explosions from the two
types of progenitors, which we quantify and discuss.
}

@@

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel to $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP''
 described by~\cite{VladWeirRyzh06},
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\citep{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning.
{\color{blue}
The flame speeds are obtained via linear interpolation within a 
three-dimensional table that combines the results of \citet{timmes92} 
and \citet{Chametal08}, and the dimensions of the table are \C{12} 
mass fraction, \Ne{22} mass fraction and log of density. While the
WD contains additional species, we have shown that treating the
abundance of \Ne{22} as a proxy for neutron-rich elements captures the
speed-up of laminar flames due to neutronization and thus reasonably 
produces variations in flame speed due to composition~\citep{jacketal2010}.}
{\color{blue} Using these tables, which were made for mixtures without $^{20}$Ne,
effectively assumes that the substitution of $^{20}$Ne for $^{16}$O will not
change the flame speeds significantly.  We consider this a reasonable
approximation given that, in the interior of our star, $^{16}$O is still much
more abundant than $^{20}$Ne, and that the flame propagation is largely
dominated by turbulent effects.}
These laminar flame speeds are boosted to account for the speed-up 
of burning due to unresolved buoyancy and background turbulence~\citep{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

@@

The burning model also describes the detonation phase with progress variables.
In this case, the model evolves thermally-activated burning with the actual
temperature-dependent rate for C consumption, which allows a propagating shock to
trigger burning, i.e.\ to propagate a detonation front. The propagating detonation
is able to describe the same stages of C burning as the deflagration case, including
the relaxation into NSE~\cite[and references therein]{townetal2016}. Finally,
we again note that the burning model was adapted for the case of burning in
hybrid white dwarfs.
{\color{blue} Parameter studies of detonations in C-O-Ne material and details of how the burning model is used to capture C-O-Ne burning may be found in \citet{willcoxetal2016}.}

@@

The hybrid white dwarf model that served as the initial condition
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included larger nuclear reaction networks than previous studies and 
thus resulted in a electron-to-baryon gradient that became unstable 
to mixing as the interior cooled, leading to a lower central C 
fraction than previous hybrid models~\citep{brooksetal2017}. We 
selected the 1.09 \Msun\ WD model 
with the 0.4 \Msun\ C/O core from \citep{brooksetal2017}.
In order to grow this WD towards the Chandrasekhar mass, it was inserted into 
a binary with a 1.4 \Msun\ He star donor having a 3 hour orbital period, similar to 
the simulations in \citet{brooksetal2016},
and the system was
allowed to evolve.  The WD grows, experiences central carbon ignition and 
simmering, and then is stopped when its central temperature reaches 
$\log T_{\rm c}(\Kelvin) = 8.9$, 
{\color{blue} 
at which point the 
temperature is just about high enough to ignite the deflagration that is 
the first stage of the explosion. The central \C{12} mass fraction 
of the model is 0.1419.}

Following the approach of \citet{willcoxetal2016}, we constructed a
corresponding ``classic'' C-O model to allow comparison of explosion
results between these hybrid models and those of previous studies
\citep{Krueger2010On-Variations-o,Kruegeretal2012}. The C-O model was 
constructed to have conditions as similar a possible to the hybrid
model (e.g.\  the two shared the same central temperature and density).  
{\color{blue} 
The central \C{12} mass fraction of this model is 0.4.
}
Figure~\ref{fig:init_conds} shows the density and temperature
profiles of the two initial one-dimensional models.
\begin{figure}
\includegraphics[width=\columnwidth]{figures/temp_dens_abund_profile.pdf}
\caption{\label{fig:init_conds}
Radial temperature, density, and  composition profiles of the one-dimensional hybrid
progenitor WD. The central temperature and density were the same as
the traditional C-O progenitor model.
The peak temperature of the hybrid model is at the center of the star,
so both the hybrid and C-O models could share the same central ignition initial conditions in our simulations.
}
\end{figure}
{\color{blue} While the C-O-Ne progenitor is the result of an evolutionary calculation,
the C-O progenitor is constructed by integrating the equation of hydrostatic equilibrium
using the Helmholtz stellar equation of state of \citet{timmes.swesty:accuracy} starting
from the central density and temperature of the C-O-Ne WD model.
This parameterized construction is reflected in the simplified
outer thermal structure for the C-O WD model appearing in Figure~\ref{fig:init_conds}.}
We centrally ignite our explosion simulations for both C-O and hybrid
models to coincide with the peak temperature at the center of the WDs,
unlike the hybrid models of~\citet{willcoxetal2016}.

@@

The simulations performed for this study consisted of a suite of
30 two-dimensional simulations of thermonuclear supernova explosions 
from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These
were compared to a suite of supernova simulations from traditional C-O 
progenitors. {\color{blue} The C-O progenitors are parameterized and 
include the effects of convective ``simmering" in the core as the WD 
approaches the Chandrasekhar mass~\citep{Chametal08,PiroBild08,jacketal2010}.
Using parameterized C-O models allowed us to choose conditions, e.g.\
central density, to control differences between the hybrid and traditional 
models and thus assess the effect of the hybrid structure.
The central density at ignition depends on both the carbon abundance 
and the binary history, mainly cooling time before accretion 
\citep{Kruegeretal2012}, and so this comparison does not correspond 
to a comparable binary scenario, but is intended as a more 
straightforward comparison.
}

@@

The ignition of the deflagration via a match head followed the
initialization described in~\citet{Kruegeretal2012}, which
followed the method of~\citet{townetal2009}.
In both the hybrid and traditional cases, the match head had a
nominal radius of $150$~km before
a different randomly-seeded perturbation was applied to the
match head for each of the 30 simulations in both suites,
The perturbation to the sphere's surface is a set of
spherical harmonic functions with randomly chosen amplitudes, and
each set of perturbations is referred to as a ``realization."
Both suites use the same 30 realizations of the ignition geometry.
These perturbations have been shown to reproduce
the scatter in \Ni{56}
yield from \SNeIa\ \citep{townetal2009}.
{\color{blue} We note that the ignition points of the hybrid models of 
\citet{willcoxetal2016} substantially differed from both the
(mixed) hybrid models presented here and  traditional models. Those progenitors
had the highest temperatures and hence were ignited at a radius of
about 300 km.}

@@

We frame the presentation of the results of the suites of simulations
principally in terms of the yield of \Ni{56}, the energy source of
the light curve of an event. \Ni{56} thus serves as a proxy
for the brightness of an explosion and comparison of the yields
is equivalent to comparing the brightness of events. The
yields were estimated from $Y_e$ and the NSE progress variable, by 
assuming the composition upon NSE freeze-out is \Ni{56} plus equal
parts \Fe{54} and \Ni{58} \citep{townetal2009,Meaketal09}.
This assumption allows the fraction of IGEs in the form of \Ni{56} to be 
estimated from the $Y_e$ tracked by the burning model. 
{\color{blue}
This process has been shown to provide estimated \Ni{56} yields consistent with
the results of detailed network and NSE calculations~\citep{townetal2016,caldertownsley2018}.
}

@@

The production of \Ni{56}, after NSE freeze-out, as a function of simulation time
for 10 simulations of explosions from each progenitor is shown in Figure \ref{fig:nithybrid}.
{\color{blue} The sharp increase in the yield occurring after approximately 1.0 second indicates
the transition to the detonation phase with its significantly faster burning. The curves
indicate that on average, the C-O-Ne models reach the DDT later then the hybrid model, 
implying more expansion of the WD and lower density burning in the detonation phase.}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/RatioNi56IGE_v_time_plot.pdf}
\caption{\label{fig:compare_ratio}
Expected ratio of \Ni{56} to total IGE production by mass vs.\ time showing the evolution
for both suites of simulations. 
{\color{blue} Shown are realizations 21-30.}
Results from C-O-Ne progenitors are in blue and C-O progenitors in red.
%SAME AS DON'S FIGURE 19
}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{figures/MBTI_v_time_plot.pdf}
\caption{\label{fig:compare_burned}
Plot of estimated mass burned to IGEs vs.\ time showing the evolution
for both suites of simulations for realizations 25-30. Results from C-O-Ne progenitors are
in blue and C-O progenitors red.
%SAME AS DON'S FIGURE 20
}
\end{figure}

{\color{blue} 
The evolution of the fraction, by mass, of IGE material expected to be in the form of \Ni{56} after 
freeze-out is shown in Figure \ref{fig:compare_ratio}. The figure presents the results from the
same realizations as in Figure \ref{fig:nithybrid}. The ratios for the yields of 
explosions from hybrid progenitors are consistently lower than the ratios of yields from
traditional progenitors.
Figure~\ref{fig:compare_burned} shows the 
concurrent evolution of mass burned to IGE for realizations 25-30 from both suites.
The total IGE has a mostly similar evolution, and the hybrid results show
more scatter, but there is not an obvious difference between the results from
the two progenitors. These plots together show that the 
ratio is the quantity that demonstrates important differences between 
C-O and C-O-Ne cases.
}

>>>>

\subsection{Proposed Progenitor Settings}\label{sec:progenitors}

@@

In the DDT paradigm, the duration of the deflagration phase sets
the amount of expansion of the star prior to the bulk of burning,
which is critical to the composition of material synthesized in the 
explosion. Also, early burning during the deflagration phase is at 
high enough densities that the effects of electron capture are 
significant and similarly influence the composition of the material
synthesized in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.

@@

Discerning the role played by hybrid progenitors in the global gamut of
thermonuclear supernovae is the goal of the study we present in this paper. Our simulations
provide the yields of \Ni{56} in the supernova explosions. Because
the radioactive decay of \Ni{56} powers the light curve of an event, 
the yield of \Ni{56} serves as a proxy for brightness of the event so
our estimates allows us to discern trends in brightness. Our study thus 
can offer insight into the relative brightness of explosions from these
progenitors as well as the overall scatter in the brightness
of observed events.

Our simulation methodology is described in the next section and
our results include a comparison of nucleosynthetic explosion yields between these
new mixed hybrid C-O-Ne progenitors and traditional C-O progenitors
in the section following the methodology. Our results show systematic
differences in the yield of \Ni{56} between explosions from the two
types of progenitors, which we quantify and discuss.

@@

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel to $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP''
 described by~\cite{VladWeirRyzh06},
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\citep{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning.
The flame speeds are obtained via linear interpolation within a 
three-dimensional table that combines the results of \citet{timmes92} 
and \citet{Chametal08}, and the dimensions of the table are \C{12} 
mass fraction, \Ne{22} mass fraction and log of density. While the
WD contains additional species, we have shown that treating the
abundance of \Ne{22} as a proxy for neutron-rich elements captures the
speed-up of laminar flames due to neutronization and thus reasonably 
produces variations in flame speed due to composition~\citep{jacketal2010}.
Using these tables, which were made for mixtures without $^{20}$Ne,
effectively assumes that the substitution of $^{20}$Ne for $^{16}$O will not
change the flame speeds significantly.  We consider this a reasonable
approximation given that, in the interior of our star, $^{16}$O is still much
more abundant than $^{20}$Ne, and that the flame propagation is largely
dominated by turbulent effects.
These laminar flame speeds are boosted to account for the speed-up 
of burning due to unresolved buoyancy and background turbulence~\citep{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

@@

The burning model also describes the detonation phase with progress variables.
In this case, the model evolves thermally-activated burning with the actual
temperature-dependent rate for C consumption, which allows a propagating shock to
trigger burning, i.e.\ to propagate a detonation front. The propagating detonation
is able to describe the same stages of C burning as the deflagration case, including
the relaxation into NSE~\cite[and references therein]{townetal2016}. Finally,
we again note that the burning model was adapted for the case of burning in
hybrid white dwarfs.
Parameter studies of detonations in C-O-Ne material and details of how the burning model 
is used to capture C-O-Ne burning may be found in \citet{willcoxetal2016}.

@@

The hybrid white dwarf model that served as the initial condition
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included larger nuclear reaction networks than previous studies and 
thus resulted in a electron-to-baryon gradient that became unstable 
to mixing as the interior cooled, leading to a lower central C 
fraction than previous hybrid models~\citep{brooksetal2017}. We 
selected the 1.09 \Msun\ WD model 
with the 0.4 \Msun\ C/O core from \citep{brooksetal2017}.
In order to grow this WD towards the Chandrasekhar mass, it was inserted into 
a binary with a 1.4 \Msun\ He star donor having a 3 hour orbital period, similar to 
the simulations in \citet{brooksetal2016},
and the system was
allowed to evolve.  The WD grows, experiences central carbon ignition and 
simmering, and then is stopped when its central temperature reaches 
$\log T_{\rm c}(\Kelvin) = 8.9$, 
at which point the 
temperature is just about high enough to ignite the deflagration that is 
the first stage of the explosion. The central \C{12} mass fraction 
of the model is 0.1419.

Following the approach of \citet{willcoxetal2016}, we constructed a
corresponding ``classic'' C-O model to allow comparison of explosion
results between these hybrid models and those of previous studies
\citep{Krueger2010On-Variations-o,Kruegeretal2012}. The C-O model was 
constructed to have conditions as similar a possible to the hybrid
model (e.g.\  the two shared the same central temperature and density).  
The central \C{12} mass fraction of this model is 0.4.
Figure~\ref{fig:init_conds} shows the density and temperature
profiles of the two initial one-dimensional models.
\begin{figure}
\includegraphics[width=\columnwidth]{figures/temp_dens_abund_profile.pdf}
\caption{\label{fig:init_conds}
Radial temperature, density, and  composition profiles of the one-dimensional hybrid
progenitor WD. The central temperature and density were the same as
the traditional C-O progenitor model.
The peak temperature of the hybrid model is at the center of the star,
so both the hybrid and C-O models could share the same central ignition initial conditions in our simulations.
}
\end{figure}
While the C-O-Ne progenitor is the result of an evolutionary calculation,
the C-O progenitor is constructed by integrating the equation of hydrostatic equilibrium
using the Helmholtz stellar equation of state of \citet{timmes.swesty:accuracy} starting
from the central density and temperature of the C-O-Ne WD model.
This parameterized construction is reflected in the simplified
outer thermal structure for the C-O WD model appearing in Figure~\ref{fig:init_conds}.
We centrally ignite our explosion simulations for both C-O and hybrid
models to coincide with the peak temperature at the center of the WDs,
unlike the hybrid models of~\citet{willcoxetal2016}.

@@

The simulations performed for this study consisted of a suite of
30 two-dimensional simulations of thermonuclear supernova explosions 
from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These
were compared to a suite of supernova simulations from traditional C-O 
progenitors. The C-O progenitors are parameterized and 
include the effects of convective ``simmering" in the core as the WD 
approaches the Chandrasekhar mass~\citep{Chametal08,PiroBild08,jacketal2010}.
Using parameterized C-O models allowed us to choose conditions, e.g.\
central density, to control differences between the hybrid and traditional 
models and thus assess the effect of the hybrid structure.
The central density at ignition depends on both the carbon abundance 
and the binary history, mainly cooling time before accretion 
\citep{Kruegeretal2012}, and so this comparison does not correspond 
to a comparable binary scenario, but is intended as a more 
straightforward comparison.

@@

The ignition of the deflagration via a match head followed the
initialization described in~\citet{Kruegeretal2012}, which
followed the method of~\citet{townetal2009}.
In both the hybrid and traditional cases, the match head had a
nominal radius of $150$~km before
a different randomly-seeded perturbation was applied to the
match head for each of the 30 simulations in both suites,
The perturbation to the sphere's surface is a set of
spherical harmonic functions with randomly chosen amplitudes, and
each set of perturbations is referred to as a ``realization."
Both suites use the same 30 realizations of the ignition geometry.
These perturbations have been shown to reproduce
the scatter in \Ni{56}
yield from \SNeIa\ \citep{townetal2009}.
We note that the ignition points of the hybrid models of 
\citet{willcoxetal2016} substantially differed from both the
(mixed) hybrid models presented here and  traditional models. Those progenitors
had the highest temperatures and hence were ignited at a radius of
about 300 km.

@@

We frame the presentation of the results of the suites of simulations
principally in terms of the yield of \Ni{56}, the energy source of
the light curve of an event. \Ni{56} thus serves as a proxy
for the brightness of an explosion and comparison of the yields
is equivalent to comparing the brightness of events. The
yields were estimated from $Y_e$ and the NSE progress variable, by 
assuming the composition upon NSE freeze-out is \Ni{56} plus equal
parts \Fe{54} and \Ni{58} \citep{townetal2009,Meaketal09}.
This assumption allows the fraction of IGEs in the form of \Ni{56} to be 
estimated from the $Y_e$ tracked by the burning model. 
This process has been shown to provide estimated \Ni{56} yields consistent with
the results of detailed network and NSE calculations~\citep{townetal2016,caldertownsley2018}.

@@

The production of \Ni{56}, after NSE freeze-out, as a function of simulation time
for 10 simulations of explosions from each progenitor is shown in Figure \ref{fig:nithybrid}.
The sharp increase in the yield occurring after approximately 1.0 second indicates
the transition to the detonation phase with its significantly faster burning. The curves
indicate that on average, the C-O-Ne models reach the DDT later then the hybrid model, 
implying more expansion of the WD and lower density burning in the detonation phase.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/RatioNi56IGE_v_time_plot.pdf}
\caption{\label{fig:compare_ratio}
Expected ratio of \Ni{56} to total IGE production by mass vs.\ time showing the evolution
for both suites of simulations. 
Shown are realizations 21-30.}
Results from C-O-Ne progenitors are in blue and C-O progenitors in red.
%SAME AS DON'S FIGURE 19
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{figures/MBTI_v_time_plot.pdf}
\caption{\label{fig:compare_burned}
Plot of estimated mass burned to IGEs vs.\ time showing the evolution
for both suites of simulations for realizations 25-30. Results from C-O-Ne progenitors are
in blue and C-O progenitors red.
%SAME AS DON'S FIGURE 20
}
\end{figure}

The evolution of the fraction, by mass, of IGE material expected to be in the form of \Ni{56} after 
freeze-out is shown in Figure \ref{fig:compare_ratio}. The figure presents the results from the
same realizations as in Figure \ref{fig:nithybrid}. The ratios for the yields of 
explosions from hybrid progenitors are consistently lower than the ratios of yields from
traditional progenitors.
Figure~\ref{fig:compare_burned} shows the 
concurrent evolution of mass burned to IGE for realizations 25-30 from both suites.
The total IGE has a mostly similar evolution, and the hybrid results show
more scatter, but there is not an obvious difference between the results from
the two progenitors. These plots together show that the 
ratio is the quantity that demonstrates important differences between 
C-O and C-O-Ne cases.
